---
title: "WUP_04"
author: "Judy Insunza"
date: "10/3/2025"
output:
  html_document:
    df_print: paged
---

```{r}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv.csv <- read_csv("GSScsv.csv")
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

This data is numerical and I will be using the helper file to solve this assignment.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
## Descriptive Results
### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(davisdata,aes(y=WORDSUM,x=DEGREE)) 
```
```{r}
basicCN + geom_boxplot()
```
  
### Numerical Descriptive Results  

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
favstats(WORDSUM~Sex,data=GSScsv.csv) 
```
Looking at the favstats, I noticed that the Q1 of graduate school is higher than the Q3 of both High school and LT high school, suggesting there will be a statistical difference between them.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv.csv)
summary.aov(model1)
```

When I observed the inferential results, I found that the P-value is  Q1 of graduate school is higher than the Q3 of both High school and LT high school, suggesting there will be a statistical difference between them.  

```{r}

par(mar=c(3,17,2,2))
plot(TukeyHSD(model1,conf.level=0.98),las=2,)
```

Looking at the 98% confidence level, I noticed that the only confident intervals that the zero does not pass through are the LT High school- Graduate and high school-graduate. This suggests that they are significantly different. 
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

For this assignment, I learned how to analyze numerical categorical data but with more than 2 categories. In fact, I was able to analyze 5 different categories. This data was comparing DEGREE to WORDSUM. I was able to plot graphical, numerical and inferential results. The numerical results using favstats showed me that  Q1 of graduate school is higher than the Q3 of both High school and LT high school, suggesting there will be a statistical difference between them. The inferential results also agreed to this idea. I was able to get the p-value which was about 0.000236, which told me there is a significant difference in at lest 2 categories.To be able to know these two categories, I plotted a 98% confidence level, I noticed that the only confident intervals that the zero does not pass through are the LT High school- Graduate and high school-graduate. This suggests that they are significantly different.